---
title:          "Analyzing the time-frequency content of EEG data"
author:         "Alexander Enge"
date:           "25/05/2022"
institute:      "Neuro Lab @ Humboldt-Universität zu Berlin"
classoption:    "t"
bibliography:   "template/references.bib"
nocite:         "@berger1929, @keil2022, @roach2008"
csl:            "template/apa.csl"
output:
  beamer_presentation:
    includes:
      in_header: "template/hu_template.tex"
---

```{r, setup, include=FALSE}
# Load packages
library(here)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  fig.align = "center",
  out.width = "90%"
)

# Create helper function to include figures only if the file exists
figures_dir <- here("template", "figures")
include_if_exists <- function(path, ...) {
  if (file.exists(path)) knitr::include_graphics(path, ...)
}
```

## Today

1. Why study neural oscillations?

    - Evoked vs. induced activity

2. Frequency analysis

    - Sine waves, Fourier transform

3. Time-frequency analysis

    - Morlet wavelets

4. Example workflow

    - MNE-Python style, hu-neuro-pipeline style

## Why study neural oscillations?

### Empirical data

\bigskip

```{r}
include_if_exists(here(figures_dir, "berger.png"))
```

::: columns

:::: column

### Algorithm

```{r, out.width="40%"}
include_if_exists(here(figures_dir, "circuit.pdf"))
```

::::

:::: column

### Computation

- Working memory

- Language

- Consciousness

::::

:::

## Evoked vs. induced activity

```{python, out.width="70%", results=FALSE}
import numpy as np
import matplotlib.pyplot as plt

from mne.time_frequency import tfr_array_morlet


def make_sine(times, freq, ampl, phase=0.):
    return ampl * np.sin(2 * np.pi * freq * times + phase * 2 * np.pi)


def make_windowed_sine(times, start, stop, freq, ampl, phase=0.):
    sine = make_sine(times, freq, ampl, phase)
    window = np.zeros_like(times)
    window[(times >= start) & (times <= stop)] = 1.
    return sine * window


# Define sampling interval
sampling_rate = 500.
tmax = 1.
times = np.arange(0, tmax, 1. / sampling_rate)

# Define Morlet wavelets
morlet_freqs = np.arange(6., 30., step=2.)
morlet_cycles = np.arange(3., 15., step=1.)

# Prepare empty matrices
signal_sum = np.zeros_like(times)
tfr_sum = np.zeros((len(morlet_freqs), len(times)))

# Prepare plot
n_trials = 4
fig, axs = plt.subplots(nrows=n_trials + 1, ncols=2, figsize=(7, 6))
for ix, ax in zip(range(n_trials), axs):

    # Create signal and plot as a timecourse
    noise = np.random.normal(0, 1, size=times.shape)
    sine_early = make_windowed_sine(
        times, start=0.2, stop=0.4, freq=10., ampl=10., phase=0.25 * ix)
    sine_late = make_windowed_sine(
        times, start=0.6, stop=0.8, freq=20., ampl=10.)
    signal = noise + sine_early + sine_late
    ax[0].plot(times, signal, 'r')

    # Add trial label
    trial_label = f'Trial {ix + 1}'
    ax[0].text(-0.02, max(signal) - 5, trial_label)

    # Compute time-frequency decomposition and plot as image
    signal_mne = np.expand_dims(signal, axis=0)
    signal_mne = np.expand_dims(signal_mne, axis=0)
    tfr = tfr_array_morlet(signal_mne, sampling_rate, freqs=morlet_freqs,
                           n_cycles=morlet_cycles, output='power')
    tfr = np.squeeze(tfr)
    ax[1].imshow(
        tfr, cmap='turbo', aspect='auto', vmin=-tfr.max(), vmax=tfr.max(),
        origin='lower', extent=(0., 1., 6., 40.))

    # Add to running sums for later averaging
    signal_sum += signal
    tfr_sum += tfr

# Plot averaged timecourse
signal_avg = signal_sum / n_trials
axs[-1, 0].plot(times, signal_avg, 'r')
axs[-1, 0].text(-0.02, max(signal_avg) - 4, 'Average')

# Plot averaged power
tfr_avg = tfr_sum / n_trials
_ = axs[-1, 1].imshow(tfr_avg, cmap='turbo', aspect='auto',
                      vmin=-tfr_avg.max(),
                      vmax=tfr_avg.max(),
                      origin='lower', extent=(0., 1., 6., 40.))

# Column titles
_ = axs[0, 0].set_title('Time domain', weight='bold')
_ = axs[0, 1].set_title('Time-frequency domain', weight='bold')

# x-axis lables
_ = axs[-1, 0].set_xlabel('Time (s)')
_ = axs[-1, 1].set_xlabel('Time (s)')

# y-axis labels
_ = axs[2, 0].set_ylabel('Amplitude (µV)')
_ = axs[2, 1].set_ylabel('Frequency (Hz)')

# Show figure
_ = fig.tight_layout()
_ = plt.show()
```

## Frequency analysis

**Goal:** Examine which frequencies (oscillations) contribute to a stretch of continuous EEG

**Approach:** Decompose the continuous EEG into a set of sine waves $\rightarrow$ Fourier transform

## Sine waves

```{python}
import matplotlib.patches as patches

# Define sine wave
freq = 2.
ampl = 5.
x = ampl * np.sin(2 * np.pi * freq * times)

# Create empty figure
fig, ax = plt.subplots(figsize=(9, 4))

# Annotate one cycle with shading
rect = patches.Rectangle(
  (0, -ampl), 1. / freq, 2 * ampl, color='r', linewidth=0, alpha=.1)
_ = plt.text(
  0.01, 0.5 - ampl, "Cycle", ha='left', va="center", color='r', alpha=.6)
_ = ax.add_patch(rect)

# Plot sine wave
_ = ax.plot(times, x, 'r')
_ = ax.set_xlabel('Time (s)')
_ = ax.set_ylabel('Amplitude (µV)')

# Annotate amplitude
_ = plt.text(0.25 * tmax / freq + 0.02, ampl / 2, "Amplitude (µV)",
         ha="center", va="center", rotation = "vertical")
_ = plt.text(0.25 * tmax / freq - 0.02, ampl / 2, "How strong?",
        ha = "center", va="center", rotation = "vertical")
_ = plt.annotate('', xy=(0.25 * tmax / freq, ampl), xytext=(0.25 * tmax / freq, 0),
             arrowprops=dict(arrowstyle='<->'))

# Annotate frequency
_ = plt.text(0.75 * tmax / freq, -ampl / 10., "Frequency (Hz)",
             ha = "center", va = "center")
_ = plt.text(0.75 * tmax / freq, ampl / 12., "How fast?",
             ha='center', va="center")
_ = plt.annotate('', xy=(0, 0), xytext=(tmax / freq, 0),
             arrowprops=dict(arrowstyle='<->'))

# Annotate phase
_ = plt.annotate('Where?\nPhase\n(°)', xy=(0, 0.5), xytext=(0, ampl),
             ha='center', va='top', arrowprops=dict(arrowstyle='->'))
```

## Sine waves

```{python, out.width="100%"}
# Plot new sine waves
smaller_ampl = 2.5
x_smaller_ampl = smaller_ampl * np.sin(2 * np.pi * freq * times)
faster_freq = 4.
x_faster_freq = ampl * np.sin(2 * np.pi * faster_freq * times)
x_shift_phase = ampl * np.sin(2 * np.pi * freq * times - np.pi / 2)
x_news = {'Amplitude 2.5 vs. 5 µV': x_smaller_ampl,
          'Frequency 4 vs. 2 Hz': x_faster_freq,
          'Phase shift 90 vs. 0°': x_shift_phase}

# Plot together with the original sine wave
fig, axs = plt.subplots(1, 3, figsize=(10, 2.5))
for ax, (title, x_new) in zip(axs, x_news.items()):
  _ = ax.plot(times, x, 'r', alpha=0.2)
  _ = ax.plot(times, x_new, 'r')
  _ = ax.set_title(title, weight='bold')

# Axis labels
_ = axs[1].set_xlabel('Time (s)')
_ = axs[0].set_ylabel('Amplitude (µV)')

# Show figure
_ = fig.tight_layout()
_ = plt.show()
```

## Fourier transform

::: columns

:::: {.column width="30%"}

```{r}
include_if_exists(here(figures_dir, "fourier.jpg"))
```

::::

:::: {.column width="70%"}

**Joseph Fourier** (1768--1830):

\bigskip

*Any signal can be expressed as sum of weighted sine waves, each with its own frequency, amplitude, and phase*

::::

:::

\scriptsize
\begin{equation*}
\begin{split}
f(t) &= A_0 + A_1 \cos(\omega t + \varphi_1) + A_2 \cos(\omega t + \varphi_2) + ... + A_N \cos(\omega t +\varphi_N) \\
&= \sum_{n=0}^N A_n \cos(\omega t +\varphi_n)\quad\text{where $A$ = amplitude, $\omega$ = frequency, $\varphi$ = phase}
\end{split}
\end{equation*}

## Fourier transform

```{python}
from scipy.signal import periodogram

# Define frequencies of interest and their amplitudes
freqs = [2., 10., 25.]
ampls = [5., 4., 2.]

# Create figure
fig, all_axs = plt.subplots(len(freqs) + 1, 2, figsize=(9, 6), sharey='col')
signal_sum = np.zeros_like(times)
for row_ix, (freq, ampl, axs) in enumerate(zip(freqs, ampls, all_axs)):

    # Create a single sine wave
    signal = ampl * np.sin(2 * np.pi * freq * times)
    _ = axs[0].plot(times, signal, 'r')

    # Annotate
    label = f'$\omega = {int(freq)}$ Hz, $A = {int(ampl)}$ µV'
    _ = axs[0].text(-0.02, 8., label)

    # Compute spectrum
    freqs, spect = periodogram(signal, sampling_rate, scaling='spectrum')
    _ = axs[1].plot(freqs, spect, 'b')
    _ = axs[1].set_xlim(0., 30.)

    # Add sine wave to the combined signal
    signal_sum += signal

# Plot combined signal
_ = all_axs[-1, 0].plot(times, signal_sum, 'r')
_ = all_axs[-1, 0].text(-0.02, 8., 'Sum')

# Plot spectrum of the combined signal
freqs, spect = periodogram(signal_sum, sampling_rate, scaling='spectrum')
_ = all_axs[-1, 1].plot(freqs, spect, 'b')
_ = all_axs[-1, 1].set_xlim(0., 30.)

# Column titles
_ = all_axs[0, 0].set_title('Time domain', weight='bold')
_ = all_axs[0, 1].set_title('Frequency domain', weight='bold')

# Axis labels
_ = all_axs[-1, 0].set_xlabel('Time (s)')
_ = all_axs[-1, 1].set_xlabel('Frequency (Hz)')
_ = fig.text(0, 0.5, 'Amplitude (µV)', va='center', rotation='vertical')
_ = fig.text(0.5, 0.5, 'Power (µV²)', va='center', rotation='vertical')

# Show figure
_ = plt.subplots_adjust(wspace=3.5)
_ = fig.tight_layout()
_ = plt.show()
```

## Frequency analysis in MNE-Python

```{python}
from mne import set_log_level

set_log_level('ERROR')
```

```{python, eval=FALSE}
# Download raw data
from os import environ, getcwd
from pipeline.datasets import ucap
environ['PIPELINE_DATA_DIR'] = getcwd() + '/data'
_ = ucap.get_paths(n_participants = 1)
```

```{python, echo=TRUE, out.width="80%"}
# Read raw data
from mne.io import read_raw_brainvision
raw = read_raw_brainvision('data/raw/05.vhdr')

# Downsample to make subsequent computations faster
raw = raw.resample(125.)

# Plot spectrum for a single channel
_ = raw.plot_psd(picks='Cz', color='b', spatial_colors=False)
```

```{python}
plt.show()
```

## Frequency analysis in MNE-Python

```{python, echo=TRUE, out.width="80%"}
# Plot spectrum for all channels, restricted to 0-30 Hz
_ = raw.plot_psd(fmax=30, color='b', spatial_colors=False)
```

```{python}
plt.show()
```

## Frequency analysis

```{python, out.width="90%", results=FALSE}
# Prepare empty matrices
signal_sum = np.zeros_like(times)
spect_sum = np.zeros((int(len(times) / 2 + 1),))
tfr_sum = np.zeros((len(morlet_freqs), len(times)))

# Prepare plot
n_trials = 4
fig, axs = plt.subplots(nrows=n_trials + 1, ncols=3, figsize=(9, 6))
for ix, ax in zip(range(n_trials), axs):

    # Create signal and plot as a timecourse
    noise = np.random.normal(0, 1, size=times.shape)
    sine_early = make_windowed_sine(
        times, start=0.2, stop=0.4, freq=10., ampl=10., phase=0.25 * ix)
    sine_late = make_windowed_sine(
        times, start=0.6, stop=0.8, freq=20., ampl=10.)
    signal = noise + sine_early + sine_late
    ax[0].plot(times, signal, 'r')

    # Add trial label
    trial_label = f'Trial {ix + 1}'
    ax[0].text(-0.02, max(signal) - 5, trial_label)

    # Compute Fourier transform and plot as a spectrum
    freqs, spect = periodogram(signal, sampling_rate, scaling='spectrum')
    _ = ax[1].plot(freqs, spect, 'b')
    _ = ax[1].set_xlim(0., 30.)

    # Compute time-frequency decomposition and plot as image
    signal_mne = np.expand_dims(signal, axis=0)
    signal_mne = np.expand_dims(signal_mne, axis=0)
    tfr = tfr_array_morlet(signal_mne, sampling_rate, freqs=morlet_freqs,
                           n_cycles=morlet_cycles, output='power')
    tfr = np.squeeze(tfr)
    ax[2].imshow(
        tfr, cmap='turbo', aspect='auto', vmin=-tfr.max(), vmax=tfr.max(),
        origin='lower', extent=(0., 1., 6., 40.))

    # Add to running sums for later averaging
    signal_sum += signal
    spect_sum += spect
    tfr_sum += tfr

# Plot averaged timecourse
signal_avg = signal_sum / n_trials
axs[-1, 0].plot(times, signal_avg, 'r')
axs[-1, 0].text(-0.02, max(signal_avg) - 4, 'Average')

# Plot averaged spectrum
spect_avg = spect_sum / n_trials
axs[-1, 1].plot(freqs, spect_avg, 'b')
axs[-1, 1].set_xlim(0., 30.)

# Plot averaged power
tfr_avg = tfr_sum / n_trials
_ = axs[-1, 2].imshow(tfr_avg, cmap='turbo', aspect='auto',
                      vmin=-tfr_avg.max(),
                      vmax=tfr_avg.max(),
                      origin='lower', extent=(0., 1., 6., 40.))

# Column titles
_ = axs[0, 0].set_title('Time domain', weight='bold')
_ = axs[0, 1].set_title('Frequency domain', weight='bold')
_ = axs[0, 2].set_title('Time-frequency domain', weight='bold')

# x-axis lables
_ = axs[-1, 0].set_xlabel('Time (s)')
_ = axs[-1, 1].set_xlabel('Frequency (Hz)')
_ = axs[-1, 2].set_xlabel('Time (s)')

# y-axis labels
_ = axs[2, 0].set_ylabel('Amplitude (µV)')
_ = axs[2, 1].set_ylabel('Power (µV²)')
_ = axs[2, 2].set_ylabel('Frequency (Hz)')

# Show figure
_ = fig.tight_layout()
_ = plt.show()
```

## Time-frequency analysis

**Goal:** Estimate how power (or phase) at each frequency changes over time, e.g., in response to a stimulus

**Approach:** Apply the Fourier transform to a short time window and move this window over time points in the epoch $\rightarrow$ Short-term Fourier transform (STFT), Morlet wavelet convolution

## Morlet wavelets

```{python, out.width="90%", results=FALSE}
from mne import events_from_annotations, Epochs
from mne.time_frequency import morlet, tfr_morlet

# Define wavelets
freqs = np.arange(6., 30., step=2.)
n_cycles = np.arange(3., 15., step=1.)

# Compute ERP example
events, _ = events_from_annotations(raw, regexp='Stimulus')
triggers = [201, 205]
epochs = Epochs(raw, events, triggers, tmin=-0.5, tmax=1.5, baseline=(-0.2, 0.))
ave = epochs.copy().load_data().pick_channels(['Cz']).average().data.squeeze()

# Compute TFR example
tfr = tfr_morlet(epochs, freqs, n_cycles, return_itc=False, average=False)
tfr = tfr.apply_baseline((None, None), mode='percent')
tfr = tfr.apply_baseline((-0.45, -0.05), mode='mean')
tfr_ave = tfr.average().copy().pick_channels(['Cz']).data.squeeze()

# Prepare multi-panel figure
fig = plt.figure(figsize=(10, 7))
widths = [10, 1, 10, 1, 10]
gs = fig.add_gridspec(len(freqs), 5, width_ratios=widths)

# Plot wavelets
wavelet_params = zip(reversed(freqs), reversed(n_cycles))
for ix, (freq, this_n_cycles) in enumerate(wavelet_params):
    wavelet = morlet(sampling_rate, [freq], [this_n_cycles])[0]
    sigma_t = this_n_cycles / (2.0 * np.pi * freq)
    times_wavelet = np.arange(0., 5. * sigma_t, 1.0 / sampling_rate)
    times_wavelet = np.r_[-times_wavelet[::-1], times_wavelet[1:]]
    _ = ax = fig.add_subplot(gs[ix, 0])
    _ = ax.plot(times_wavelet, wavelet, "r")
    if ix == 0:
        _ = ax.set_title('Wavelets', weight='bold')
    if ix == len(freqs) - 1:
        _ = ax.set_xlabel('Time (s)')
_ = fig.text(0.06, 0.5, 'Amplitude (µV)', va='center', rotation='vertical')

# Add convolution symbol
ax = fig.add_subplot(gs[:, 1])
_ = ax.text(-1., 0.47, '$*$')
_ = plt.axis('off')

# Plot ERP time course
ax = fig.add_subplot(gs[5:7, 2])
_ = ax.plot(epochs.times, ave * 1e6, 'r')
_ = ax.set_title('EEG timecourse', weight='bold')
_ = ax.set_xlabel('Time (s)')
_ = ax.set_ylabel('Amplitude (µV)')

# Add equals symbol
ax = fig.add_subplot(gs[:, 3])
_ = ax.text(-1., 0.47, '$=$')
_ = plt.axis('off')

# Plot time-frequency image
ax = fig.add_subplot(gs[5:7, 4])
im = ax.imshow(
    tfr_ave, cmap='turbo', aspect='auto', vmin=-tfr_ave.max(),
    vmax=tfr_ave.max(), origin='lower',
    extent=(min(epochs.times), max(epochs.times), min(freqs), max(freqs)))
_ = ax.set_title('Time-frequency plot', weight='bold')
cbar = plt.colorbar(im)
_ = cbar.set_label('Power', rotation=-90)
_ = ax.set_xlabel('Time (s)')
_ = ax.set_ylabel('Frequency (Hz)')

# Show figure
_ = plt.tight_layout()
_ = plt.subplots_adjust(hspace=0., wspace=0.3)
_ = plt.show()
```

## Baseline correction

- Should fit at least one full cycle at the lowest frequency

    - E.g., 5 Hz $\rightarrow$ min. 200 ms baseline

- Should account for the $1/f$ scaling $\rightarrow$ Divisive baseline

    - Subtract + divide by mean baseline: Percent signal change

    - Divide by mean baseline + take logarithm: Decibel

- Baseline should end *before* rather than *at* stimulus onset

    - Prevents post-stimulus activity "smearing" into the baseline

## Baseline correction - UPDATE

**Problem:** Divisive single-trial pre-stimulus baseline correction creates positive bias in single-trial post-stimulus power [@grandchamp2011; @hu2014]

**Solution:**

- Apply a first, *divisive* baseline using the *entire epoch* as the baseline window

- Then apply a second, *subtractive* baseline using the *pre-stimulus interval only*

- Now implemented in the hu-neuro-pipeline, with options for the divisive baseline method (`tfr_baseline_mode`, e.g., percent, ratio, $z$-score) and for the subtractive baseline window (`tfr_baseline_tmin`, `tfr_baseline_tmax`)

## Time-frequency analysis in MNE-Python

```{python, echo=TRUE}
# Load functions
import numpy as np
from mne import events_from_annotations, Epochs
from mne.time_frequency import tfr_morlet

# Segment continuous signal to epochs
events, _ = events_from_annotations(raw, regexp='Stimulus')
triggers = [201, 205]
epochs = Epochs(raw, events, triggers, tmin=-0.5, tmax=1.5, baseline=(-0.2, 0.))
print(epochs.get_data().shape) # Dimensions: (trials, channels, time points)
```

```{python, echo=TRUE}
# Apply Morlet wavelet decomposition
freqs = np.arange(6., 30., step=2.)
n_cycles = np.arange(3., 15., step=1.)
tfr = tfr_morlet(epochs, freqs, n_cycles, return_itc=False, average=False)
print(tfr.data.shape) # Dimensions: (trials, channels, frequencies, time points)
```

## Time-frequency analysis in MNE-Python

```{python, echo=TRUE, results=FALSE}
# Divisive baseline correction to get percent signal change
tfr = tfr.apply_baseline((None, None), mode='percent')

# Subtractive baseline correction using the pre-stimulus interval only
tfr = tfr.apply_baseline((-0.45, -0.05), mode='mean')

# Plot power at one channel, averaged across epochs
tfr_ave = tfr.average()
tfr_ave.plot(picks='Cz', cmap='turbo')
```

```{python, out.width="60%"}
_ = plt.show()
```

## Time-frequency analysis in MNE-Python

```{python, echo=TRUE, results=FALSE}
# Compute power per condition and the difference between conditions
tfr_aves = {'Blurred': tfr['201'].average(),
            'Normal': tfr['205'].average(),
            'Difference': tfr['201'].average() - tfr['205'].average()}

# Create multi-panel plot
import matplotlib.pyplot as plt
fig, axs = plt.subplots(nrows=1, ncols=len(tfr_aves), figsize=(9, 2))
for ix, (label, tfr_ave) in enumerate(tfr_aves.items()):
    axs[ix].set_title(label, weight='bold')
    tfr_ave.plot(picks='Cz', cmap='turbo', vmin=-0.7, vmax=0.7,
                 axes=axs[ix], show=False)
```

```{python, out.width="95%"}
_ = plt.tight_layout()
_ = plt.show()
```

## Time-frequency analysis with the pipeline

General settings:

```{r, echo=TRUE, eval=FALSE}
# Import the Python package from R
pipeline <- reticulate::import("pipeline")

# Run the pipeline with the `tfr` (time-frequency) options
res <- pipeline$group_pipeline(
  vhdr_files = "data/raw",
  log_files = "data/log",
  output_dir = "output",
  ocular_correction = "data/cali",
  triggers = c(201:208, 211:218),
  average_by = c("n_b", "DeviantPosRL", "n_b/DeviantPosRL"),
  perform_tfr = TRUE,
  tfr_freqs = seq(6, 30, by = 2),
  tfr_cycles = seq(3, 15, by = 1),
  tfr_baseline_tmin = -0.45,
  tfr_baseline_tmax = -0.05,
  tfr_baseline_mode = "percent",
  ... # See next 2 slides
)
```

## Time-frequency analysis with the pipeline

For effects with *a priori* knowledge about their distribution:

```{r, echo=TRUE, eval=FALSE}
# Define time-frequency "components" of interest
res <- pipeline$group_pipeline(
  ...,
  tfr_components = list(
    "name" = list("alpha"),
    "tmin" = list(0.0),
    "tmax" = list(0.2),
    "fmin" = list(8),
    "fmax" = list(14),
    "roi" = list(
      c("PO9", "PO7", "PO3", "POz", "PO4", "PO8", "PO10", "O1", "Oz", "O2")
    )
  )
)
```

$\rightarrow$ Then use the single trial data frame to run mixed models

## Time-frequency analysis with the pipeline

For exploratory analyses $\rightarrow$ Cluster-based permutation tests:

```{r, echo=TRUE, eval=FALSE}
# Run 3D cluster-based permutation tests for contrasts of interest
res <- pipeline$group_pipeline(
  ...,
  perm_contrasts = list(
    c("blurr", "normal"),
    c("blurr/re", "blurr/li"),
    c("normal/re", "normal/li")
  )
)
```

$\rightarrow$ Creates a new data frame with cluster-level $p$-values across time, space (channels), and frequencies

## General remarks

- Richer view of the EEG signal (evoked + induced activity)

- Many new parameters to think carefully about:

    - Design $\rightarrow$ Longer inter-trial interval; jittering

    - Wavelet frequencies + number of cycles $\rightarrow$ T/f tradeoff

    - Baseline correction window and method

    - Interpretation:
        
        - Narrow-band vs. broad-band
        
        - Oscillation vs. rate of change

- Everything matters

## Thanks

## References

\scriptsize
